Master Thesis

### 2013-06-16 the-last-part.tex
Finished the graphical ACG chapter

$$
N_{aucun} : N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S)
$$
$$
N_{le_1} : N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S)
$$
$$
N_{le_2} : N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S)
$$

@@

$$
N_{ne_{tv_1}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S)
$$
$$
N_{ne_{tv_2}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
$$

@@

$$
N_{ne_{tv_3}} : (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}T \limp S)
$$

@@

$$
N_{que_1} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F
$$
$$
N_{que_2} : (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T
$$
$$
N_{de_1} : NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F
$$
$$
N_{de_2} : NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T
$$
$$
N_{de_3} : NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
$$

@@

$$
E_{dort_1} : NP\_VAR{=}F \limp S\_EXT{=}NO
$$
$$
E_{dort_2} : NP\_VAR{=}T \limp S\_EXT{=}ROOT
$$
$$
E_{aime_1} : NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO
$$
$$
E_{aime_2} : NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT
$$
$$
E_{aime_3} : NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT
$$
$$
E_{dit\ que_1} : NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO
$$
$$
E_{dit\ que_2} : NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT
$$
$$
E_{dit\ que_3} : NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB
$$
$$
E_{dit\ que_4} : NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
$$

@@

$$
E_{qui} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
$$
$$
E_{que_1} : (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N
$$
$$
E_{que_2} : (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
$$

@@

$$
E_{le_1} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}NO) \limp S\_EXT{=}NO)
$$
$$
E_{le_2} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}ROOT) \limp S\_EXT{=}ROOT)
$$
$$
E_{le_3} : N \limp ((NP\_VAR{=}F \limp S\_EXT{=}EMB) \limp S\_EXT{=}EMB)
$$

@@

$$
E_{de} : NP\_VAR{=}F \limp N \limp N
$$

@@

As was the case with our treatment of negation, we refined our familiar
types by tacking on new kinds of information and defined rules for how
these pieces of information percolate up the parse tree all the way to
the topmost level still pertinent for ensuring grammaticality. We will
now look at another similar example.

\section{Agreement}

Finally, we will cover the phenomenon of number agreement. This
motivates a straightforward refinement of our types, where all noun and
noun phrase types carry number information.

$$
A_{Marie} : NP\_NUM{=}SG
$$
$$
A_{tatou} : N\_NUM{=}SG
$$
$$
A_{tatous} : N\_NUM{=}PL
$$
$$
A_{le} : N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S)
$$
$$
A_{les} : N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S)
$$
$$
A_{dort} : NP\_NUM{=}SG \limp S
$$
$$
A_{dorment} : NP\_NUM{=}PL \limp S
$$
$$
A_{aime_1} : NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S
$$
$$
A_{aime_2} : NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S
$$
$$
A_{aiment_1} : NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S
$$
$$
A_{aiment_2} : NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S
$$
$$
A_{qui_1} : (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG
$$
$$
A_{qui_2} : (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL
$$
$$
A_{de_1} : NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG
$$
$$
A_{de_2} : NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG
$$
$$
A_{de_3} : NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL
$$
$$
A_{de_4} : NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
$$

@@

In the next section, we will see what it takes to enforce all of the
constraints above (negation, extraction and agreement) at the same time,
which will give us some idea on how to go about building a grammar which
handles a wide variety of such phenomena.

\section{Putting It All Together}

@@

$$
C_{de_1} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_2} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_3} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_4} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_5} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_6} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_7} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_8} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_9} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_{10}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
$$
$$
C_{de_{11}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG})
$$
$$
C_{de_{12}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
$$

@@

$$
\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in \{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}
$$
$$
C_{de_{1(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y})
$$
$$
C_{de_{2(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
$$
$$
C_{de_{3(\textcolor{blue}{x}, \textcolor{blue}{y})}} : (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
$$

@@

$$
\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in\{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}, \forall \textcolor{red}{p}, \textcolor{red}{q} \in \{\textcolor{red}{F}, \textcolor{red}{T}\}, \textcolor{red}{p \land q} \neq \textcolor{red}{T}
$$
$$
C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} : (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
$$

@@

Furthermore, the current notation does not even make it obvious that
these are independent mechanisms. One has to examine the type to make
sure that there is no interaction. In the next two subsections, we will
devote ourselves to investigating ways of disentangling these.

@@

\section{Intersections of Languages}

We will now turn to another solution to the complexity of grammars
implementing multiple independent mechanisms.

When ACGs are discussed, one of the oft-praised properties is
compositionality. Let us consider a situation where we have three
signatures $\Sigma_1$, $\Sigma_2$ and $\Sigma_3$, their respective
distinguished types $S_1$, $S_2$ and $S_3$ and two lexicons
$\mathcal{L}_{12} : \Sigma_1 \to \Sigma_2$ and $\mathcal{L}_{23} :
\Sigma_2 \to \Sigma_3$ such that $\mathcal{L}_{12}(S_1) = S_2$ and
$\mathcal{L}_{23}(S_2) = S_3$. (TODO: INCLUDE A DIAGRAM OF THIS
SITUATION)

@@

Further than that, we can look at the languages $\mathcal{L}_{12}(A_1)$
and $\mathcal{L}_{23}(A_2)$, images of the abstract languages given by
the lexicons. Since the lexicons are homomorphic mappings w.r.t. the
types, all elements of $\mathcal{L}_{12}(A_1)$ will have the type
$\mathcal{L}_{12}(S_1) = S_2$ and will therefore also belong to $A_2$
(similarly for $\mathcal{L}_{23}(A_2)$ and $A_3$). These new languages,
which are formed by mapping an ACG abstract language using a lexicon,
are called \emph{object languages}. Their terms are constrained not only
by the types of the signature they are built upon but also by the type
system of the abstract signature.

Finally, we can also take the composition of lexicons $\mathcal{L}_{13}
= \mathcal{L}_{23} \circ \mathcal{L}_{12}$, which itself is also a
lexicon, and map it over $A_1$. This gives us a language
$\mathcal{L}_{13}(A_1) = \mathcal{L}_{23}(\mathcal{L}_{12}(A_1))$ which
is at the same time the image of the object language
$\mathcal{L}_{12}(A_1)$ given by $\mathcal{L}_{23}$ and also the image
of the abstract language $A_1$ given by $\mathcal{L}_{13}$. This points
to an interesting property of ACG languages: they are closed under
transformation by a lexicon.

@@

The definition of ACGs given above is quite practical in that it is
complete, it gives a precise account of what set of languages are
generated by ACGs (abstract and object languages for which there exist
the quadruples described above). This contrasts with the presentation of
ACGs that we gave in~\ref{sec:acg}, where we defined ACGs as
collections of signatures with associated distinguished types and
connected with lexicons. Our definition only introduced the requisite
machinery, but delayed the delimitation of the way languages can be
defined. Leaving this question open will let us answer it now by
providing an extension/generalization of ACGs which will help us
overcome the modularity issues that have been the theme of this
chapter.\footnote{Incidentally, this was not the reason for the way we
  presented ACGs in~\ref{sec:acg}. We chose this style as it grouped
  together the information in a more natural way. We approach signatures
  as descriptions of domains (strings, syntax, semantics), each one
  being associated with a distinguished type telling us which objects
  the domain has set out to describe (strings, sentences, truth
  values). Having distinguished types directly attached to signatures
  allows us to omit them from the ACG quadruples. However, this only
  leaves us with triples containing the two signatures and the lexicon
  connecting them, and since we can infer the signatures themselves from
  the domain and the range of the lexicon, we do not have much
  motivation to introduce these tuples. Instead of composing ACGs, we
  compose just the lexicons.

On top of that, we will want to define signatures and lexicons that
yield different object languages (at least some string representation of
the form and some semantic representation of the meaning), which would
mean having at least two different ACGs which are linked (e.g. by
sharing the same abstract signature). However, since these two grammars
are very closely related, we would prefer to reason about our grammar as
\emph{one} system/graph of signatures and lexicons which defines
multiple languages instead of talking about \emph{multiple} grammars
which share some similarities.

>>>>

\begin{align*}
N_{aucun} &: N\_NEG{=}F \limp ((NP\_NEG{=}T \limp S) \limp S) \\
N_{le_1} &: N\_NEG{=}F \limp ((NP\_NEG{=}F \limp S) \limp S) \\
N_{le_2} &: N\_NEG{=}T \limp ((NP\_NEG{=}T \limp S) \limp S) \\
\end{align*}

@@

\begin{align*}
N_{ne_{tv_1}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}F \limp S) \\
N_{ne_{tv_2}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}F \limp NP\_NEG{=}T \limp S)
\end{align*}

@@

\begin{align*}
N_{ne_{tv_3}} &: (NP\_NEG{=}F \limp NP\_NEG{=}F \limp S) \limp (NP\_NEG{=}T \limp NP\_NEG{=}T \limp S)
\end{align*}

@@

\begin{align*}
N_{que_1} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}F \limp N\_NEG{=}F \\
N_{que_2} &: (NP\_NEG{=}F \limp S) \limp N\_NEG{=}T \limp N\_NEG{=}T \\
N_{de_1} &: NP\_NEG{=}F \limp N\_NEG{=}F \limp N\_NEG{=}F \\
N_{de_2} &: NP\_NEG{=}F \limp N\_NEG{=}T \limp N\_NEG{=}T \\
N_{de_3} &: NP\_NEG{=}T \limp N\_NEG{=}F \limp N\_NEG{=}T
\end{align*}

@@

\begin{align*}
E_{dort_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \\
E_{dort_2} &: NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
E_{aime_1} &: NP\_VAR{=}F \limp NP\_VAR{=}F \limp S\_EXT{=}NO \\
E_{aime_2} &: NP\_VAR{=}T \limp NP\_VAR{=}F \limp S\_EXT{=}ROOT \\
E_{aime_3} &: NP\_VAR{=}F \limp NP\_VAR{=}T \limp S\_EXT{=}ROOT \\
E_{dit\ que_1} &: NP\_VAR{=}F \limp S\_EXT{=}NO \limp S\_EXT{=}NO \\
E_{dit\ que_2} &: NP\_VAR{=}T \limp S\_EXT{=}NO \limp S\_EXT{=}ROOT \\
E_{dit\ que_3} &: NP\_VAR{=}F \limp S\_EXT{=}ROOT \limp S\_EXT{=}EMB \\
E_{dit\ que_4} &: NP\_VAR{=}F \limp S\_EXT{=}EMB \limp S\_EXT{=}EMB
\end{align*}

@@

\begin{align*}
E_{qui} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
E_{que_1} &: (NP\_VAR{=}T \limp S\_EXT{=}ROOT) \limp N \limp N \\
E_{que_2} &: (NP\_VAR{=}T \limp S\_EXT{=}EMB) \limp N \limp N
\end{align*}

@@

\begin{align*}
E_{le_1} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}NO) \limp S\_EXT{=}NO) \\
E_{le_2} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}ROOT) \limp S\_EXT{=}ROOT) \\
E_{le_3} &: N \limp ((NP\_VAR{=}F \limp S\_EXT{=}EMB) \limp S\_EXT{=}EMB)
\end{align*}

@@

\begin{align*}
E_{de} &: NP\_VAR{=}F \limp N \limp N
\end{align*}

@@

As was the case with our treatment of negation, we refined our familiar
types by tacking on new kinds of information and defined rules for how
these pieces of information percolate up the parse tree all the way to
the topmost level still pertinent for ensuring grammaticality.

\section{Agreement}
\label{sec:agreement}

Finally, we will look at another example of implementing a linguistic
constraint in ACGs by covering the phenomenon of number
agreement.

Number agreement motivates a straightforward refinement of our types,
where all noun and noun phrase types carry number information.

\begin{align*}
A_{Marie} &: NP\_NUM{=}SG \\
A_{tatou} &: N\_NUM{=}SG \\
A_{tatous} &: N\_NUM{=}PL \\
A_{le} &: N\_NUM{=}SG \limp ((NP\_NUM{=}SG \limp S) \limp S) \\
A_{les} &: N\_NUM{=}PL \limp ((NP\_NUM{=}PL \limp S) \limp S) \\
A_{dort} &: NP\_NUM{=}SG \limp S \\
A_{dorment} &: NP\_NUM{=}PL \limp S \\
A_{aime_1} &: NP\_NUM{=}SG \limp NP\_NUM{=}SG \limp S \\
A_{aime_2} &: NP\_NUM{=}SG \limp NP\_NUM{=}PL \limp S \\
A_{aiment_1} &: NP\_NUM{=}PL \limp NP\_NUM{=}SG \limp S \\
A_{aiment_2} &: NP\_NUM{=}PL \limp NP\_NUM{=}PL \limp S \\
A_{qui_1} &: (NP\_NUM{=}SG \limp S) \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
A_{qui_2} &: (NP\_NUM{=}PL \limp S) \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
A_{de_1} &: NP\_NUM{=}SG \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
A_{de_2} &: NP\_NUM{=}PL \limp N\_NUM{=}SG \limp N\_NUM{=}SG \\
A_{de_3} &: NP\_NUM{=}SG \limp N\_NUM{=}PL \limp N\_NUM{=}PL \\
A_{de_4} &: NP\_NUM{=}PL \limp N\_Pl \limp N\_NUM{=}PL
\end{align*}

@@

\section{Putting It All Together}
\label{sec:together}

In this section, we will see what it takes to enforce all of the
constraints introduced in the preceding sections (negation, extraction
and agreement) at the same time, which will give us some idea on how to
go about building a grammar which handles a wide variety of such
phenomena.

@@

\begin{align*}
C_{de_1} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_2} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_3} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_4} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_5} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_6} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_7} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_8} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_9} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_{10}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL}) \\
C_{de_{11}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}SG}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}SG}) \\
C_{de_{12}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}PL}) \limp
(N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}PL})
\end{align*}

@@

$$
\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in \{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}
$$
\begin{align*}
C_{de_{1(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \\
C_{de_{2(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}F}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y}) \\
C_{de_{3(\textcolor{blue}{x}, \textcolor{blue}{y})}} &: (NP\_\textcolor{red}{NEG{=}T}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}F}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}T}\_\textcolor{blue}{NUM{=}y})
\end{align*}

@@

$$
\forall \textcolor{blue}{x}, \textcolor{blue}{y} \in\{\textcolor{blue}{SG}, \textcolor{blue}{PL}\}, \forall \textcolor{red}{p}, \textcolor{red}{q} \in \{\textcolor{red}{F}, \textcolor{red}{T}\}, \textcolor{red}{p \land q} \neq \textcolor{red}{T}
$$
\begin{align*}
C_{de(\textcolor{blue}{x}, \textcolor{blue}{y}, \textcolor{red}{p}, \textcolor{red}{q})} &: (NP\_\textcolor{red}{NEG{=}p}\_\textcolor{green}{VAR{=}F}\_\textcolor{blue}{NUM{=}x}) \limp (N\_\textcolor{red}{NEG{=}q}\_\textcolor{blue}{NUM{=}y}) \limp (N\_\textcolor{red}{NEG{=}(p \lor q)}\_\textcolor{blue}{NUM{=}y})
\end{align*}

@@

Furthermore, the current notation does not even make it obvious that
these are independent mechanisms. One has to examine the type to make
sure that there is no interaction. In the rest of this chapter, we will
devote ourselves to investigating ways of disentangling these.

@@

\section{Intersections of Languages}
\label{sec:sects-of-langs}

We will turn to another solution to the problem of complex grammars that
try to implement multiple independent constraints. This solution will be
based on a generalization of ACGs, but in order to introduce this
generalization, we will first need to present the manner in which ACGs
and their compositions are formulated in ACG literature.

\subsection{The Usual Formalization of ACGs}
\label{ssec:usual-acgs}

Let us consider a system with three signatures $\Sigma_1$, $\Sigma_2$
and $\Sigma_3$, their respective distinguished types $S_1$, $S_2$ and
$S_3$ and two lexicons $\mathcal{L}_{12} : \Sigma_1 \to \Sigma_2$ and
$\mathcal{L}_{23} : \Sigma_2 \to \Sigma_3$ such that
$\mathcal{L}_{12}(S_1) = S_2$ and $\mathcal{L}_{23}(S_2) = S_3$. See
Figure~\ref{fig:acg-serial-comp} for a diagrammatic visualization of
this structure.

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/acg-serial-comp.pdf}
    \caption{\label{fig:acg-serial-comp} Serial composition of ACGs.}
  \end{subfigure}
  \qquad
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[height=0.2\textheight]{diagrams/acg-parallel-comp.pdf}
    \caption{\label{fig:acg-parallel-comp} Parallel composition of ACGs.}
  \end{subfigure}
  \caption{\label{fig:acg-comp-modes} Two modes of composition for
    ACGs.}
\end{figure}

@@

Further than that, we can look at the languages $\mathcal{L}_{12}(A_1)$
and $\mathcal{L}_{23}(A_2)$, images of the abstract languages given by
the lexicons. Since the lexicons are homomorphic mappings w.r.t. the
types, all elements of $\mathcal{L}_{12}(A_1)$ will have the type
$\mathcal{L}_{12}(S_1) = S_2$ and will therefore also belong to $A_2$
(similarly, we have that $\mathcal{L}_{23}(A_2) \subseteq A_3$). These
new languages, which are formed by mapping an ACG abstract language
using a lexicon, are called \emph{object languages}. Their terms are
constrained not only by the types of the signature they are built upon
but also by the type system of the abstract signature.

Finally, we can also take the composition of lexicons $\mathcal{L}_{13}
= \mathcal{L}_{23} \circ \mathcal{L}_{12}$, which itself is also a
lexicon, and map it over $A_1$. This gives us a language
$\mathcal{L}_{13}(A_1) = \mathcal{L}_{23}(\mathcal{L}_{12}(A_1))$ which
is at the same time the image of the object language
$\mathcal{L}_{12}(A_1)$ given by $\mathcal{L}_{23}$ and also the image
of the abstract language $A_1$ given by $\mathcal{L}_{13}$ (thus it is
itself also an object language). This points to an interesting property
of ACG languages: they are closed under transformation by a lexicon.

@@

The definition of ACGs given above is quite practical in that it is
complete, i.e. it gives a precise account of what set of languages are
generated by ACGs (abstract and object languages for which there exist
the quadruples described above). This contrasts with the presentation of
ACGs that we gave in~\ref{sec:acg}, where we defined ACGs as collections
of signatures associated with distinguished types and connected with
lexicons. Our definition only introduced the requisite machinery, but
delayed the delimitation of the way languages can be defined. Leaving
this question open will let us answer it now by providing an extension
of ACGs which will help us overcome the modularity issues that have been
the theme of this chapter.\footnote{Incidentally, this was not the
  reason for the way we presented ACGs in~\ref{sec:acg}. We chose this
  style as it grouped together the information in a more natural way. We
  approach signatures as descriptions of domains (strings, syntax,
  semantics), each one being associated with a distinguished type
  telling us which objects the domain has set out to describe (strings,
  sentences, truth values). Having distinguished types directly attached
  to signatures allows us to omit them from the ACG quadruples,
  $\mathopen{<}\Sigma_A, \Sigma_O, \mathcal{L},
  \cancel{S}\mathclose{>}$. However, this only leaves us with triples
  containing the two signatures and the lexicon connecting them, and
  since we can infer the signatures themselves from the domain and the
  range of the lexicon $\mathcal{L} : \Sigma_A \to \Sigma_O$, we do not
  have much motivation to introduce these tuples,
  $\mathopen{<}\cancel{\Sigma_A}, \cancel{\Sigma_O}, \mathcal{L},
  \cancel{S}\mathclose{>}$. Instead of composing ACGs, we compose just
  the lexicons.
